---
title: "01 Peak Fitting Data Reduction"
author: "Tristan Caro"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup

## Libraries
```{r}
library(tidyverse)
library(readxl)
library(openxlsx)
library(magrittr)
```

## Function definitions

```{r}
select_fit_params <- function(df) {
  # the deparse(substitute(x)) changes the argument into a string that can be used
  sample_id_str = deparse(substitute(df))
  df %>% 
    select(fit, p, a , g, w) %>% 
    # add cell ids:
    mutate(sample = sample_id_str)
}

select_fit_areas <- function(df) {
  # the deparse(substitute(x)) changes the argument into a string that can be used
  sample_id_str = deparse(substitute(df))
  df %>% 
    # there are two "cell" columns - pick the first one
    select(unique(colnames(.))) %>% 
    select(cell, CD, contains("CH"), `CD%`) %>% 
    mutate(sample = sample_id_str) %>% 
    # remove NA rows introduced from fit dataset
    filter(!is.na(cell))
}

select_fit_xy <- function(df) {
  # change the argument into a string that will be used later
  sample_id_str = deparse(substitute(df))
  df %>% 
    select(unique(colnames(.))) %>% 
    # exclude columns from the other datasets
    select(-c(contains("CH"), contains("Peak2"), CD, `CD%`, p, a , g, w, fit)) %>% 
    # tidy it! pivot the dataset in order to change column names to x values
    pivot_longer(cols = -c(cell),
      values_to = "intensity",
      names_to = "wavenumber_cm") %>%
    # change these values to numeric, if not already
    mutate(
      wavenumber_cm = as.numeric(wavenumber_cm),
      intensity = as.numeric(intensity)
    ) %>% 
    mutate(sample = sample_id_str)
}
```

# Read data

```{r}
# Define where our large xlsx file is
path <- "data/Raman/fitted_data/CD_2F_RamanFits_FINAL.xlsx"

# Use openxlsx to get the names of the sheets
sheets <- openxlsx::getSheetNames(path)

# Read in all the data by lapply-ing the read.xlsx function across all the sheet names
fitted_data <- lapply(sheets, openxlsx::read.xlsx, xlsxFile=path)

# Assign names to the data frame
names(fitted_data) <- sheets
```

# Reduce Data

## Pull excel sheets into their own dataframes

I could probably do this more elegantly, but for now this will suffice.

```{r}
Thy_00_07 <- fitted_data$`Thy_00_07!`
Thy_10_08 <- fitted_data$`Thy_10_08!`
Thy_20_09 <- fitted_data$`Thy_20_09!`
Thy_30_10 <- fitted_data$`Thy_30_10!`
Thy_40_11 <- fitted_data$`Thy_40_11!`
Thy_50_12 <- fitted_data$`Thy_50_12!`

Mb_00_13 <- fitted_data$`Mb_00_13!`
Mb_10_14 <- fitted_data$`Mb_10_14!`
Mb_20_23 <- fitted_data$`Mb_20_23!`
Mb_30_27 <- fitted_data$`Mb_30_27!`
Mb_40_y <- fitted_data$`Mb_40_y!`
Mb_50_18 <- fitted_data$`Mb_50_18!`

```

## Separate the data

Here we separate the fit parameters, peak area summaries, and fitted peak paths (xy data) into their own dataframes, identifying them by the sample name.
```{r}
fit_data <- bind_rows(
  select_fit_params(Thy_00_07),
  select_fit_params(Thy_10_08),
  select_fit_params(Thy_20_09),
  select_fit_params(Thy_30_10),
  select_fit_params(Thy_40_11),
  select_fit_params(Thy_50_12),
  select_fit_params(Mb_00_13),
  select_fit_params(Mb_10_14),
  select_fit_params(Mb_20_23),
  select_fit_params(Mb_30_27),
  select_fit_params(Mb_40_y),
  select_fit_params(Mb_50_18)
)

area_data <- bind_rows(
  select_fit_areas(Thy_00_07),
  select_fit_areas(Thy_10_08),
  select_fit_areas(Thy_20_09),
  select_fit_areas(Thy_30_10),
  select_fit_areas(Thy_40_11),
  select_fit_areas(Thy_50_12),
  select_fit_areas(Mb_00_13),
  select_fit_areas(Mb_10_14),
  select_fit_areas(Mb_20_23),
  select_fit_areas(Mb_30_27),
  select_fit_areas(Mb_40_y),
  select_fit_areas(Mb_50_18)
) %>% 
  mutate(
    cell_id = paste0(
      sample,
      "_", 
      sprintf("%02d", cell) # keep the cell id digits in 2 digits!
      ),
    f_l = case_when(
      sample == "Mb_40_y" ~ 40,
      sample == "Mb_20_x" ~ 20,
      TRUE ~ as.numeric(
        str_sub(
          cell_id, 
          start = -8, 
          end = -7)
        )
      )
    ) %>% 
  # remove the cell column
  select(-cell) %>% 
  # reorder the columns
  relocate(sample, cell_id)



peak_data_xy <- bind_rows(
  select_fit_xy(Thy_00_07),
  select_fit_xy(Thy_10_08),
  select_fit_xy(Thy_20_09),
  select_fit_xy(Thy_30_10),
  select_fit_xy(Thy_40_11),
  select_fit_xy(Thy_50_12),
  select_fit_xy(Mb_00_13),
  select_fit_xy(Mb_10_14),
  select_fit_xy(Mb_20_23),
  select_fit_xy(Mb_30_27),
  select_fit_xy(Mb_40_y),
  select_fit_xy(Mb_50_18)
) %>% 
  mutate(
    cell_id = paste0(
      sample,
      "_", 
      sprintf("%02d", cell) # keep the cell id digits in 2 digits!
      )
    ) %>% 
  # remove the cell column
  select(-cell) %>% 
  # reorder the columns
  relocate(sample, cell_id)
```


# Clear memory

```{r}
rm(
  Thy_00_07,
  Thy_10_08,
  Thy_20_09,
  Thy_30_10,
  Thy_40_11,
  Thy_50_12,
  Mb_00_13,
  Mb_10_14,
  Mb_20_23,
  Mb_30_27,
  Mb_40_y,
  Mb_50_18
  )
```


# Export

```{r}
saveRDS(peak_data_xy, file = "cache/peak_data_xy.RDS")
saveRDS(area_data, file = "cache/area_data.RDS")
saveRDS(fit_data, file = "cache/fit_data.RDS")
```



